---
title: "allometric_growth_analysis"
output: html_document
---

#Descripton
This script performs an allometric analysis of coral growth using buoyant weight measurements. It filters and reshapes initial and final weights, merges in tank metadata, tests for initial mass differences across treatments, fits an allometric mixed-effects model(log-log), and evaluates treatment effects (fish and wound) on allometrically scaled growth using likelihood ratio tests.

# Setup
```{r}
# ===============================
# 1. Load Required Libraries
# ===============================
library(tidyverse)
library(janitor)
library(here)
library(gt)
library(broom.mixed)
library(ggpubr)
library(lme4)

#...........................aesthetics...........................
wound_palette <- c("Large" = "#FF8B00", 
                   "Small" = "#be8333", 
                   "No Wound" = "#485900")

fish_palette <- c("Fish" = "#2b4b52", 
                  "No Fish" = "#ad301a")
# ===============================
# 2. Load Buoyant Weight Data
# ===============================
bw_raw <- read_csv(here("data", "fish_regen_buoyantweight (1).csv")) %>%
  clean_names() %>%
  filter(coral_id != 45)%>%
  filter(coral_id != 91)

# Split initial and final weights
bw_initial <- bw_raw %>%
  filter(date == "initial") %>%
  select(coral_id, wound, fish, initial_weight = bouyantweight_g)

bw_final <- bw_raw %>%
  filter(date == "final") %>%
  select(coral_id, final_weight = bouyantweight_g)

# Join initial and final to get one row per coral
bw_wide <- left_join(bw_initial, bw_final, by = "coral_id") %>%
  mutate(
    delta_mass = final_weight - initial_weight,
    coral_id = as.character(coral_id)
  ) 



  #.....................Load Surface Area Data.....................
sa <- read_csv(here("data", "fish_regen_mastersheet_wound closure_necrosis_sa - mastersheet.csv")) %>% 
  clean_names() %>% 
  mutate(coral_id = as.character(coral_id)) %>% 
  select(-fish, -date_collected, -date_fragment) %>% 
  mutate(sa_col=ifelse(wound == "Large", sa_cal - 3, #removing 3 cm^2 from large wounds and 1 from small
                       ifelse(wound == "Small", sa_cal - 1, sa_cal))) %>% 
  select(-wound) #removing the column so it doesn't duplicate when we join with bw_wide

bw_sa_wide <- left_join(bw_wide, sa, by = "coral_id") %>% 
  mutate(
    wound = factor(wound, levels = c("No Wound", "Small", "Large")),
    tank = as.factor(tank)
  ) %>% 
  mutate(tank = factor(tank, levels = sort(unique(as.numeric(as.character(tank))))))
  
# ===============================
# 3. Load Tank Metadata
# ===============================
tank_df <- read_csv(here("data", "Copy of fish_regen_mastersheet_wound closure_necrosis - mastersheet.csv")) %>%
  clean_names() %>%
  select(coral_id, tank) %>%
  mutate(coral_id = as.character(coral_id))

# ===============================
# 4. Merge Tank Info into Wide Format
# ===============================
bw_merged <- left_join(bw_wide, tank_df, by = "coral_id") %>%
  mutate(
    wound = factor(wound, levels = c("No Wound", "Small", "Large")),
    tank = as.factor(tank)
  )

```

#Transform Data for Allometric Modeling
```{r}
# Create a dataset with log-transformed weights and growth ratio
# Drop NAs to ensure valid model fitting
log_df <- bw_merged %>%
  mutate(
    log_i_weight = log(initial_weight),
    log_f_weight = log(final_weight),
    growth_ratio = final_weight / initial_weight
  ) %>%
  drop_na(log_i_weight, log_f_weight, tank)

# ===============================
# 4. Fixed-Effects Allometric Model
# ===============================
# Fit a linear mixed-effects model:
# Predict final log-weight from initial log-weight
# Include tank as a random intercept to account for tank-level variation

model_lmer <- lmer(log_f_weight ~ log_i_weight + (1 | tank), data = log_df)

# Summary with fixed effect (slope), intercept, and random effects
summary(model_lmer)

# Extract allometric slope (b-value) with 95% confidence interval
b_value <- tidy(model_lmer, effects = "fixed", conf.int = TRUE) %>%
  filter(term == "log_i_weight")
print(b_value)


# Generate predicted log final weights across observed range of initial weights
fit_df <- log_df %>%
  select(log_i_weight) %>%
  distinct() %>%
  arrange(log_i_weight) %>%
  mutate(predicted = predict(model_lmer, newdata = ., re.form = NA))


#  Plot Allometric Fit and Observed Data
ggplot(log_df, aes(x = log_i_weight, y = log_f_weight)) +
  geom_point(size = 2.2, shape = 21, fill = "steelblue", color = "black", alpha = 0.8, stroke = 0.3) +
  geom_line(data = fit_df, aes(x = log_i_weight, y = predicted), color = "darkred", linewidth = 1.2) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "gray60", linewidth = 0.8) +
  labs(
    title = "Allometric Scaling of Coral Buoyant Weight",
    subtitle = "Mixed-effects model with tank as a random intercept",
    x = "Log Initial Weight (g)",
    y = "Log Final Weight (g)"
  ) +
  theme_pubr(base_size = 14)


library(gt)

# Extract and prepare the b-value data
b_value_gt <- tidy(model_lmer, effects = "fixed", conf.int = TRUE) %>%
  filter(term == "log_i_weight") %>%
  transmute(
    Term = "Allometric slope (b-value)",
    Estimate = round(estimate, 3),
    `95% CI` = paste0("[", round(conf.low, 3), ", ", round(conf.high, 3), "]")
  )

# Create and format the gt table for the b value 
b_value_gt %>%
  gt() %>%
  tab_header(
    title = "Estimated Allometric Slope",
    subtitle = "Mixed-effects model with tank as a random intercept"
  ) %>%
  cols_label(
    Term = "",
    Estimate = "Estimate",
    `95% CI` = "95% Confidence Interval"
  ) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )
#..........Statistical Analysis of Growth/SA/Time Metric.........

bw_sa_stats <- bw_sa_df %>%
  drop_na(initial_weight, final_weight, fish, wound, tank) %>%
  mutate(
    wound = factor(wound, levels = c("No Wound", "Small", "Large")),
    fish = factor(fish),
    tank = factor(tank),
    bw_sa_time = (final_weight - initial_weight) / (sa_cal * 21) 
  )


#visualize the relationship between delta mass and surface area, not a super strong relationsip
#i like final/initial or log(final/initial^b)

ggplot(bw_sa_stats, aes(x = sa_cal, y = delta_mass, color = wound)) +
  geom_point(size = 2.2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.1) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Coral Growth as a Function of Surface Area",
    subtitle = "Colored by Wound Treatment",
    x = "Surface Area (cm²)",
    y = expression((Final - Initial) / (Surface_Area * Time)),
    color = "Wound Treatment"
  ) +
  theme_pubr(base_size = 14)

```

#Statistical Analysis of Allometrically Scaled Growth
```{r}


# --- Step 1: Prepare Data with Allometric Scaling ---
b_est <- b_value$estimate[1]  # b-value from previous section

growth_stats <- bw_merged %>%
  drop_na(initial_weight, final_weight, fish, wound, tank) %>%
  mutate(
    wound = factor(wound, levels = c("No Wound", "Small", "Large")),
    fish = factor(fish),
    tank = factor(tank),
    growth_scaled = log(final_weight/initial_weight ^ b_est) 
  )

# --- Step 2: Fit Mixed-Effects Models ---

# Full model with interaction
mod_int <- lmer(growth_scaled ~ fish * wound + (1 | tank), data = growth_stats)

# Additive model without interaction
mod_add <- lmer(growth_scaled ~ fish + wound + (1 | tank), data = growth_stats)

# Fish-only model
mod_fish <- lmer(growth_scaled ~ fish + (1 | tank), data = growth_stats)

# Wound-only model
mod_wound <- lmer(growth_scaled ~ wound + (1 | tank), data = growth_stats)

# Null model (intercept + tank)
mod_null <- lmer(growth_scaled ~ 1 + (1 | tank), data = growth_stats)

# --- Step 3: Likelihood Ratio Tests ---
lrt_interaction <- anova(mod_add, mod_int, test = "Chisq")
lrt_fish <- anova(mod_wound, mod_add, test = "Chisq")
lrt_wound <- anova(mod_fish, mod_add, test = "Chisq")

# --- Step 4: Summarize Results in a Table ---


lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(
    lrt_interaction$Chisq[2],
    lrt_fish$Chisq[2],
    lrt_wound$Chisq[2]
  ),
  Df = c(
    lrt_interaction$Df[2],
    lrt_fish$Df[2],
    lrt_wound$Df[2]
  ),
  p_value = c(
    lrt_interaction$`Pr(>Chisq)`[2],
    lrt_fish$`Pr(>Chisq)`[2],
    lrt_wound$`Pr(>Chisq)`[2]
  )
) %>%
  mutate(across(where(is.numeric), round, 3))

# Format with gt
lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table 2. Likelihood Ratio Tests for Effects on Coral Growth",
    subtitle = paste("Scaled growth:", expression((final - initial)/initial^b))
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )

#..........Statistical Analysis of Growth/SA/Time Metric.........

# Prep data: drop NAs, ensure proper factor levels

bw_sa_df <- bw_sa_wide %>%

  drop_na(initial_weight, final_weight, wound) %>%

  filter(coral_id != 103) %>% 

  mutate(

    delta_mass = final_weight - initial_weight,

    wound = factor(wound, levels = c("No Wound", "Small", "Large"))

  )

bw_sa_stats <- bw_sa_df %>%
  drop_na(initial_weight, final_weight, fish, wound, tank) %>%
  mutate(
    wound = factor(wound, levels = c("No Wound", "Small", "Large")),
    fish = factor(fish),
    tank = factor(tank),
    bw_sa_time = (final_weight - initial_weight) / (sa_cal * 21) 
  )


#Visualize the relationship
ggplot(bw_sa_stats, aes(x = sa_cal, y = delta_mass, color = wound)) +
  geom_point(size = 2.2, alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1.1) +
  scale_color_brewer(palette = "Dark2") +
  labs(
    title = "Coral Growth as a Function of Surface Area",
    subtitle = "Colored by Wound Treatment",
    x = "Surface Area (cm²)",
    y = expression((Final - Initial) / (Surface_Area * Time)),
    color = "Wound Treatment"
  ) +
  theme_pubr(base_size = 14)

# --- Step 2: Fit Mixed-Effects Models ---

# Full model with interaction
mod_int_sa <- lmer(bw_sa_time ~ fish * wound + (1 | tank), data = bw_sa_stats)

# Additive model without interaction
mod_add_sa <- lmer(bw_sa_time ~ fish + wound + (1 | tank), data = bw_sa_stats)

# Fish-only model
mod_fish_sa <- lmer(bw_sa_time ~ fish + (1 | tank), data = bw_sa_stats)

# Wound-only model
mod_wound_sa <- lmer(bw_sa_time ~ wound + (1 | tank), data = bw_sa_stats)

# Null model (intercept + tank)
mod_null_sa <- lmer(bw_sa_time ~ 1 + (1 | tank), data = bw_sa_stats)

# --- Step 3: Likelihood Ratio Tests ---
lrt_interaction_sa <- anova(mod_add_sa, mod_int_sa, test = "Chisq")
lrt_fish_sa <- anova(mod_wound_sa, mod_add_sa, test = "Chisq")
lrt_wound_sa <- anova(mod_fish_sa, mod_add_sa, test = "Chisq")

# --- Step 4: Summarize Results in a Table ---


lrt_table <- tibble(
  Test = c("Interaction (Fish × Wound)", "Fish Effect", "Wound Effect"),
  ChiSq = c(
    lrt_interaction_sa$Chisq[2],
    lrt_fish_sa$Chisq[2],
    lrt_wound_sa$Chisq[2]
  ),
  Df = c(
    lrt_interaction_sa$Df[2],
    lrt_fish_sa$Df[2],
    lrt_wound_sa$Df[2]
  ),
  p_value = c(
    lrt_interaction_sa$`Pr(>Chisq)`[2],
    lrt_fish_sa$`Pr(>Chisq)`[2],
    lrt_wound_sa$`Pr(>Chisq)`[2]
  )
) %>%
  mutate(across(where(is.numeric), round, 3))

# Format with gt
lrt_table %>%
  gt() %>%
  tab_header(
    title = "Table 2. Likelihood Ratio Tests for Effects on Coral Growth as a function of Surface Area and Time",
    subtitle = paste("Scaled growth:", expression((final - initial)/(surface_area * 21)))
  ) %>%
  cols_label(
    Test = "Model Comparison",
    ChiSq = "Chi-squared",
    Df = "Degrees of Freedom",
    p_value = "P-value"
  ) %>%
  fmt_number(columns = vars(ChiSq, p_value), decimals = 3) %>%
  tab_options(
    table.font.size = 14,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 13
  )

```

#Visualizing growth rate by wound and fish presence (summary line plot)
```{r}

  bw_sa_stats$wound <- factor(bw_sa_stats$wound, levels = c("Large", "Small", "No Wound")) 
  bw_sa_stats$fsh <-  factor(bw_sa_stats$fish, levels = c("No Fish", "Fish"))


  growth_rate_plot<- bw_sa_stats %>% 
    ggplot(aes(x = wound, y = bw_sa_time, color = fish, shape = fish)) +
  
  geom_jitter(
    aes(color = fish),
    position = position_jitterdodge(jitter.width = 0.1, dodge.width = 0.6),
    size = 2.5,
    alpha = 0.25,
    shape = 16
  ) +
  
  # Mean points (dodged)
  stat_summary(
    fun = mean,
    geom = "point",
    size = 5,
    position = position_dodge(width = 0.6)
  ) +
  
  stat_summary(
    fun.data = function(x) {
      data.frame(
        y = mean(x),
        ymin = mean(x) - sd(x),
        ymax = mean(x) + sd(x)
      )
    },
    geom = "errorbar",
    width = 0,               
    size = 1,
    position = position_dodge(width = 0.6)
  ) +
  
  scale_color_manual(values = fish_palette) +
 scale_shape_manual(values = c(16, 18)) +

  
  theme_minimal() +
  theme(
    panel.grid = element_blank(),
    axis.title = element_text(size = 15, face = "bold"),
    axis.text.x = element_text(size = 15),
    axis.text.y = element_text(size = 10),
    legend.position = "none",
    axis.line.x = element_line(color = "black", size = 0.5),
    axis.line.y = element_line(color = "black", size = 0.5)
  ) +
  labs(
    y = "Growth Rate (mg/cm²/day)",
    x = "Wound Size"
  )

ggsave(
  filename = here("figures", "growth", "summary_line_growth_rate_by_wound_fish.pdf"),
  plot = growth_rate_plot,
  width = 7,
  height = 5,
  dpi = 300
)

```

